Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-09 latex/main.tex
Updates from Overleaf

%%% TODO: Mention RIPPLES somewhere here? Or just in Discussion?
We compared sc2ts against three state-of-the-art methods that detect inter-lineage recombinants:
(1) RecombinHunt \cite{Alfonsi2024},
(2) CovRecomb \cite{Li2024}, and
(3) rebar (https://github.com/phac-nml/rebar).
These methods use the same underlying approach: 
they scan candidate recombinant sequences for mutations associated with different Pango lineages 
without using a phylogeny.

@@

\begin{table}
\caption{
Concordance among methods in characterizing Pango X lineages. 
The Pango X lineages analyzed here are associated with the Type 1 recombination events (related to Table 1).
For RecombinHunt and CovRecomb, the concordance rates are calculated 
based on the Pango Xs for which results are available from the original studies.
Also, for RecombinHunt, two sets of detection results are available, 
one set obtained using mutational profiles computed from a global GISAID dataset and another from a Nextstrain dataset.}
\label{tab:method_concordance}
\begin{tabular}{llll}
\hline
\textbf{Method} & \textbf{Concordant in parental lineages (\%)} & \textbf{Concordant in breakpoint interval (\%)} & \textbf{Distance (bases)} \\ [1ex]
\hline
Sc2ts & 14 (93.3) & 13 (86.7) & 540.5 (315, 766) \\ [1ex]
\hline
RH-GISAID & 13 (100.0) & 10 (76.9) & 764.0 (4, 1202) \\ [1ex]
\hline
RH-Nextstrain & 12 (85.7) & 11 (78.6) & 1160.0 (559, 2785) \\ [1ex]
\hline
CovRecomb & 7 (87.5) & 4 (50.0) & 1562.5 (217, 13208) \\ [1ex]
\hline
\end{tabular}
\end{table}

@@

For CovRecomb, we summarized the detection results for the Pango Xs as follows. 
First, we obtained the independent recombination events 
in Table S4 corresponding to a Pango X, if available. 
When multiple events corresponded to a Pango X, 
we took the results of the event having the most epidemic recombinants. 
Then, for other Pango Xs, we collated the data from a file (URL) 
containing all the detection results produced by the study. 
For each of these Pango Xs, 
we took the most commonly observed combination of 
left parent, right parent, and breakpoint interval.
We adjusted the CovRecomb-identified breakpoint interval 
so that they are right-exclusive by adding one to the right coordinate.

>>>>

%%% TODO: Mention RIPPLES somewhere here? Or just in Discussion?
We compared sc2ts against three state-of-the-art methods that detect inter-lineage recombinants:
RecombinHunt \cite{Alfonsi2024},
CovRecomb \cite{Li2024}, and
rebar (https://github.com/phac-nml/rebar).
These methods use the same underlying approach:
they scan candidate recombinant sequences for mutations associated with different Pango lineages
without using a phylogeny.

@@

\begin{table}
\caption{Concordance among methods in characterizing Pango X lineages.}
\label{tab:method_concordance}
\begin{tabular}{lccc}
\toprule & \multicolumn{1}{c}{Parent lineages} & \multicolumn{2}{c}{Breakpoint intervals} \\
\cmidrule(lr){2-2} \cmidrule(lr){3-4}
Method & Concordant (\%) & Concordant (\%) & Distance (bases) \\
\midrule
Sc2ts & 14 / 15 (93.3) & 13 / 15 (86.7) & 540.5 (315, 766)\\
RecombinHunt-GISAID & 13 / 13 (100.0) & 10 / 13 (76.9) & 764.0 (4, 1202)\\
RecombinHunt-Nextstrain & 12 / 14 (85.7) & 11 / 14 (78.6) & 1160.0 (559, 2785)\\
CovRecomb & 7 / 8 (87.5) & 4 / 8 (50.0) & 1562.5 (217, 13208)\\
\end{tabular}
\end{table}

@@

For CovRecomb, we summarized the detection results for the Pango Xs as follows. 
First, we obtained the "independent" recombination events 
corresponding to the Pango Xs (if available) 
identified by Li et al.\cite{Li2024} (see Table S4 of the study).
When multiple events corresponded to a Pango X, 
we took the results of the event having the most epidemic recombinants. 
Then, for other Pango Xs, we collated the data from a file (URL) 
containing all the detection results produced by the study. 
For each of these Pango Xs, 
we took the most commonly observed combination of 
left parent, right parent, and breakpoint interval.
We adjusted the CovRecomb-identified breakpoint interval 
so that they are right-exclusive by adding one to the right coordinate.
